---
title: "Stat 850 Project Rough Draft"
author: "Drew Hultquist, Tyler Wiederich, Wyatt McLeod"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, echo = F)
options(digits = 3)
```

# Abstract

The city of Lincoln, Nebraska maintains online open access data. Several of these data sets pertain to incidents recorded by the Lincoln Fire Department. This analysis explores response times, mapping visualizations, association of weather and traffic accidents, and a breakdown of significance by criteria based on dates.

# Introduction

## Data Set

<https://opendata.lincoln.ne.gov/datasets/LincolnNE>::lfr-apparatus-2019/about

<https://opendata.lincoln.ne.gov/datasets/lfr-incidents-2019/explore>

<https://opendata.lincoln.ne.gov/maps/941bdcfcaa4b460581461537687d8a92/about> fire zone map square miles of fire zone max distanced traveled

```{r}
library(tidyverse)
incidents <- read.csv("LFR_incidents_2019_CLEAN.csv")
apparatus <- read.csv("LFR_apparatus_2019_CLEAN.csv")
apparatus2 = apparatus %>% 
  mutate(ctm_received = mdy_hms(ctm_received),
         ctm_dispatch = mdy_hms(ctm_dispatch),
         ctm_enroute = mdy_hms(ctm_enroute),
         ctm_arrive = mdy_hms(ctm_arrive),
         ctm_clear = mdy_hms(ctm_clear),
         total_commit_time = ctm_clear - ctm_received)

library(skimr)

```

```{r}
library(sf)
library(ggplot2)



```



```{r, include=F}
source('TW_source.R')
```
```{r, include=F}
source('WM_Source.R')
```
# Selected Maps

This map shows the locations of each fire station in Lincoln in 2019, without Station 15 as it was completed late in 2019. There are 14 fire zones shown on the map, and clicking on the markers shows the station name as well as its street address.

```{r base_map}
base_map
```

The following map is a visualization of the number of dispatches that each fire zone received over the course of 2019. The color scheme goes from white to navy blue, and will show the exact number of dispatches for that zone when clicked on. *(Currently the data is not placed correctly)*

```{r dispatch_map}
dispatch_map
```

Turnout time, otherwise known as the time it takes for a fire engine to leave the station from the time it receives a call, is an important metric for evaluating fire stations. The following map shows that most fire stations in Lincoln have relatively similar turnout times, all around 80 seconds.

```{r turnout_map}
turnout_map
```

Lastly, it may be useful to know about how far away in minutes a fire engine may be if it receives a call for your residence. This map (once made fully interactive) will show the distance a fire engine can cover in 3 minutes, 5 minutes and 7 minutes for any of the fire stations found in Lincoln.

```{r times}
times
```

# Correlation of Dispatches and Weather

One may think that the weather has a discernible impact on the number of times that the services of LFD are required. Rain or snow could potentially cause additional LFD resources. However, Figure \@ref(fig:plot-pairs-temp) shows that there is little correlation between the count of dispatches and temperature reports.

```{r plot-pairs-temp, fig.height = 3, fig.cap='The correlation of the count of dispatches is less than 0.2 for all temperature categories. In this plot, TMIN is the minimum daily temperature, TAVG is the average daily temperature, and TMAX is the maximum daily temperature.'}
plot_pairs_temp
```

Especially in the Count column and row of Figure \@ref(fig:plot-pairs-temp), it is clear that temperature (and hence seasonal) relationship with the number of dispatches is not correlated.

Another metric to consider is the amount of precipitation levels. Figure \@ref(fig:plot-pairs-precip) shows the relationship between precipitation and the number of dispatches per day. Measurements are averaged over multiple weather station in Lincoln.

```{r plot-pairs-precip, fig.height = 3, fig.cap='The correlation between the number of dispatches and the amount of precipitation is less than 0.04. In this plot, precip is the amount of precipitation, snow is the amount of snowfall, and snow_dpth is the amount of snow on the ground.'}
plot_pairs_precip
```

As evident in the Count column and row for Figure \@ref(fig:plot-pairs-precip), there is very little correlation between the amount of precipitation and the number of dispatches from LFD.

# Correlation of Reported Traffic Accidents and the Number of Dispatches

In the case of traffic accidents, it is reasonable to assume that an increased number of accidents would influence the number of dispatches from LFD. Bad accidents may require additional services. However, the correlation coefficient between accidents and the number of dispatches is `r round(acc_corr,3)`. Due to a low correlation, this connection will not have any further exploration.

# Breakdown of Dates and their Significance on the Number of Dispatches

The next research question deals with if any partition of the date had an impact on the number of dispatches. Through the combinations of models for weekday, month, and weekday classification (weekend or weekday), the following model holds significance:

$$y_{ijk}=\mu + w_i+m_j+\epsilon_{ijk}$$ where

$\mu=\text{Average over all months and day classifications}$

$w_i=\text{Weekday classification where 0 = Weekday and 1 = Weekend}$

$m_j=\text{The corresponding month (1 = January, 2 = February, etc.)}$

$\epsilon_{ijk} \sim N(0,\sigma^2)=\text{Error variation for each observation}$

One assumption is that the responses are independently and identically distributed. In the Appendix, it is shown that we can satisfy this assumption by checking the autocorrelations of counts for each date. Table \@ref(tab:wcat-mt-table) shows the corresponding ANOVA table for this model.

```{r wcat-mt-table}
knitr::kable(anova_wcat_mt, digits = 4,
             caption = 'The effect of month and day weekday classification are both significant')
```

Upon further inspection using Tukey's HSD, the contrasts in Table \@ref(tab:tab-t-hsd) are significant. We see that August had significantly more dispatches than March, June, and December. Additionally, we see that the difference between the number of dispatches is significantly higher for weekdays than for weekends.

```{r tab-t-hsd}
tukey_sig_wcat_mt %>% 
  select(-Term) %>% 
  knitr::kable(caption = 'Significant contrasts of the model using months and day of week category')
```

In Table \@ref(tab:tab-wcat-mt-fitted), the fitted model values are shown for each month and day of week classification.

```{r tab-wcat-mt-fitted}
knitr::kable(wcat_mt_fitted, col.names = c('Month', 'Weekday', 'Weekend'),
             caption = 'Fitted values')
```

Figure \@ref(fig:plot-wmt-heat) visually displays these differences across day of the week.

```{r plot-wmt-heat, fig.height=3, fig.cap='August has a higher average. It is also shown that multiple weedays have higher averages than weekends.'}
plot_heatmap_wday_mt + 
  scico::scale_fill_scico(palette = "lajolla",
                          direction = 1)
```
#Top 5 call types for each station
```{r} 
apparatus2 %>% 
  group_by(station, call_type) %>% 
  summarise(Count = n()) %>% 
  arrange(station, desc(Count)) %>% 
  mutate(Rank = rank(desc(Count), ties.method = 'average')) %>% 
  ungroup() %>% 
  group_by(station) %>% 
  mutate(prop = Count/sum(Count)) %>% 
  filter(Rank <= 5 & station != 98 & station != 0) %>% 
  ungroup() %>% 
  ggplot() +
  geom_point(aes(x = station, y = call_type, color = prop),
             show.legend = T, size = 5) +
  theme_bw() +
  labs(title = 'Top Five Call Types by Station',
       x = 'Station',
       y = 'Call Type',
       color = 'Count') +
  scale_x_continuous(breaks = 1:15) +
  ggthemes::scale_color_gradient_tableau('Blue-Teal') +
  theme(panel.grid.minor = element_blank())

```
We see that the main call types of each station is Medical Delta, Medical Charlie, Medical Bravo, Medical Alpha, and Fire Alpha. 

#Top 5 units used for each station
```{r} 
apparatus2 %>% 
  group_by(station, unit_descr) %>% 
  summarise(Count = n()) %>% 
  arrange(station, desc(Count)) %>% 
  mutate(Rank = rank(desc(Count), ties.method = 'average')) %>% 
  ungroup() %>% 
  group_by(station) %>% 
  mutate(prop = Count/sum(Count)) %>% 
  filter(Rank <= 5 & station != 98 & station != 0) %>% 
  ungroup() %>% 
  ggplot() +
  geom_point(aes(x = station, y = unit_descr, color = prop),
             show.legend = T, size = 5) +
  theme_bw() +
  labs(title = 'Top Five Units Used by Station',
       x = 'Station',
       y = 'Unit Description',
       color = 'Count') +
  scale_x_continuous(breaks = 1:15) +
  ggthemes::scale_color_gradient_tableau('Blue-Teal') +
  theme(panel.grid.minor = element_blank())

```
We see that many stations do not even use five units. The most common ones used are the Engine and ALS unit. 


#Top 5 incidents responded too by each station
```{r} 
apparatus2 %>% 
  group_by(station, inc_desc) %>% 
  summarise(Count = n()) %>% 
  arrange(station, desc(Count)) %>% 
  mutate(Rank = rank(desc(Count), ties.method = 'average')) %>% 
  ungroup() %>% 
  group_by(station) %>% 
  mutate(prop = Count/sum(Count)) %>% 
  filter(Rank <= 5 & station != 98 & station != 0) %>% 
  ungroup() %>% 
  ggplot() +
  geom_point(aes(x = station, y = inc_desc, color = prop),
             show.legend = T, size = 5) +
  theme_bw() +
  labs(title = 'Top Five Incidents Responded too by Station',
       x = 'Station',
       y = 'Incident Description',
       color = 'Count') +
  scale_x_continuous(breaks = 1:15) +
  ggthemes::scale_color_gradient_tableau('Blue-Teal') +
  theme(panel.grid.minor = element_blank())
```
The main incidents appears to be Rescue/EMS. The other common ones are false alarm, canceled in route, and building fire. 

#Top 5 Actions taken by each station
```{r} 
apparatus2 %>% 
  group_by(station, action_desc) %>% 
  summarise(Count = n()) %>% 
  arrange(station, desc(Count)) %>%
  filter(is.na(action_desc) == F) %>% 
  mutate(Rank = rank(desc(Count), ties.method = 'average')) %>% 
  ungroup() %>% 
  group_by(station) %>% 
  mutate(prop = Count/sum(Count)) %>% 
  filter(Rank <= 5 & station != 98 & station != 0) %>% 
  ungroup() %>% 
  ggplot() +
  geom_point(aes(x = station, y = action_desc, color = prop, size = Rank),
             show.legend = T, size = 5) +
  theme_bw() +
  labs(title = 'Top Five Actions taken by Station',
       x = 'Station',
       y = 'Action Taken',
       color = 'Count') +
  scale_x_continuous(breaks = 1:15) +
  ggthemes::scale_color_gradient_tableau('Blue-Teal') +
  theme(panel.grid.minor = element_blank())
```
The top actions taken by a fire station appear to be standby, investigate, incident command, and cancelled in route. 
#Average Response Time Per Station
```{r} 
apparatus2 %>% 
  mutate(total_responce_time = (ctm_arrive - ctm_dispatch)) %>% 
  group_by(station, total_responce_time) %>% 

  filter(station != 98  ) %>% 
  ggplot() +
  geom_point(aes(x = station, y = total_responce_time))+
  ylim(0,6500)+
             
  
  labs(title =  'Incident Response Time by Station',
       x = 'Station',
       y = 'Total Response Time (Seconds)',
       color = 'Count') +
  scale_x_continuous(breaks = 1:15) 
```
Station 15 appears to have the best (lowest) response time. 

# Appendix

## Independence of observations

One assumption for the count of dispatches per day is that each day is independent. To check this, we can verify if there is any significant autocorrelation. Figure \@ref(fig:plot-acf-dispatch) shows the autocorrelation plot using the count of daily dispatches as a time series. Although there appears to be some significant autocorrelation at lag 1 and lag 3 (probably due to the weekday category effect), these do not strongly suggest that the data is dependent on trends. Therefore we can reasonably assume

```{r plot-acf-dispatch, fit.height = 3, fig.cap='Autocorrelation plot of daily dispatch counts. This verifies that daily observations can be assumed to have independence from discernible trends'}
plot_acf_dispatch
```
